Supermassive black holes and their host spheroids \\ III. A universal $M_{\rm BH} - n_{\rm sph}$ correlation

### 2015-12-03 papers/mbhn/mbhn.tex
pum

\shorttitle{Mbh-n paper}
\shortauthors{Savorgnan}

@@

\begin{abstract}
The S\'ersic $R^{1/n}$ model is the best approximation known to date for describing the light distribution of stellar spheroidal and disk components, 
with the S\'ersic index $n$ providing a direct measure of the central radial concentration of stars. 
The S\'ersic index of a galaxy's spheroidal component, $n_{\rm sph}$, has been shown to tightly correlate 
with the mass of the central supermassive black hole, $M_{\rm BH}$. 
The $M_{\rm BH} - n_{\rm sph}$ correlation is also expected from other two well known scaling relations: 
the $L_{\rm sph} - n_{\rm sph}$ and the $M_{\rm BH} - L_{\rm sph}$, 
where $L_{\rm sph}$ indicates the galaxy's spheroid luminosity. 
However, it is widely recognized that obtaining an accurate estimate of the spheroid S\'ersic index is a challenging task, 
and -- not surprisingly -- some studies have failed to recover a statistically significant $M_{\rm BH} - n_{\rm sph}$ correlation. 
With the aim of re-investigating the $M_{\rm BH} - n_{\rm sph}$ 
and other black hole mass scaling relations, 
we performed a detailed (i.e.~bulge, disks, bars, spiral arms, rings, halo, nucleus, etc.) 
decomposition of 66 galaxies, with directly measured black hole masses, 
that had been imaged at $3.6\rm~\mu m$ with \emph{Spitzer}. 
In this paper, the third of its series, we present an analysis of the $L_{\rm sph} - n_{\rm sph}$ and $M_{\rm BH} - n_{\rm sph}$ diagrams. 
While early-type (elliptical+lenticular) and late-type (spiral) galaxies split into two separate relations 
in the $L_{\rm sph} - n_{\rm sph}$ and $M_{\rm BH} - L_{\rm sph}$ diagrams, 
they reunite in a single $M_{\rm BH} \propto n_{\rm sph}^{3.39 \pm 0.15}$ sequence
with relatively small intrinsic scatter ($\epsilon_{(Y|X)} \simeq 0.25 \rm~dex$). 
{\bf to be finished...}
\end{abstract}

@@

and the relation between the central black hole mass, $M_{\rm BH}$, and the spheroid luminosity (e.g.~\citealt{magorrian1998}), 

@@

Motivated by the need of more accurate galaxy decompositions to refine and re-investigate scaling relations 

@@

\section{Analysis and results}
\label{sec:anal}
As in \emph{Paper II}, a linear regression analysis of the $L_{\rm sph} - n_{\rm sph}$ 
(Table \ref{tab:lregLn} and Figure \ref{fig:magn})
and $M_{\rm BH} - n_{\rm sph}$ (Table \ref{tab:lregMn} and Figure \ref{fig:mbhn}) diagrams 
was performed using three different routines: 
the BCES code from \cite{akritasbershady1996}, 
the FITEXY routine \citep{press1992}, as modified by \cite{tremaine2002}, 
and the Bayesian estimator {\tt linmix\_err} \citep{linmixerr}.
All of these three routines take into account the intrinsic scatter, 
but only the FITEXY and the {\tt linmix\_err} codes allow one to quantify it.
\cite{tremaine2002} cautioned that 
the BCES estimator becomes ineffective when the dataset contains at least one low-precision measurement -- 
regardless of how many high-precision measurements are in the sample -- 
and tends to be biased in case of low number statistics 
or if the mean square of the uncertainties associated to the independent variable is comparable to 
the variance of the distribution of the independent variable. 
According to the results from the Monte Carlo Markov Chain simulations of \cite{tremaine2002} and \cite{novak2006}, 
these problems can be overcome with the use of the modified FITEXY routine. 
\cite{park2012} also concluded that the modified FITEXY routine performs better and returns less biased results than the BCES estimator, 
and noted that the the modified FITEXY routine is computationally less intensive than the Bayesian technique {\tt linmix\_err}. 
Given that at least one of our subsamples (the lenticular galaxies) has a small size 
and that the uncertainties associated to $n_{\rm sph}$ are relatively large 
compared to the range spanned by the $n_{\rm sph}$ values for most of our subsamples, 
we put more trust in the results obtained with the modified FITEXY routine 
and throughout the text we quote only those. \\ 
We report both symmetrical and nonsymmetrical linear regressions.  
Symmetrical regressions are meant to be compared with theoretical expectations, 
whereas nonsymmetrical forward ($Y|X$) regressions -- which minimize the scatter in the $Y$ direction -- 
allow one to predict the value of the observable $Y$ with the best possible precision. \\
We searched for extreme outliers in both the $L_{\rm sph} - n_{\rm sph}$ and $M_{\rm BH} - n_{\rm sph}$ diagrams, 
and found that in our $L_{\rm sph} - n_{\rm sph}$ plot there are no $3\sigma$ outliers, 
whereas in our $M_{\rm BH} - n_{\rm sph}$ plot 
the lenticular galaxies NGC 0524 and NGC 3998 reside more than $3\sigma$ from the bisector linear regression for all galaxies. 
These two galaxies have therefore been excluded from the rest of the analysis. \\

@@

\section{Conclusions}

@@

%%
\acknowledgments
kkkkk

>>>>

\shorttitle{A universal $M_{\rm BH} - n_{\rm sph}$ correlation}
\shortauthors{Savorgnan}

@@

\begin{abstract}
The S\'ersic $R^{1/n}$ model is the best approximation known to date for describing the light distribution of stellar spheroidal and disk components, 
with the S\'ersic index $n$ providing a direct measure of the central radial concentration of stars. 
The S\'ersic index of a galaxy's spheroidal component, $n_{\rm sph}$, has been shown to tightly correlate 
with the mass of the central supermassive black hole, $M_{\rm BH}$. 
The $M_{\rm BH} - n_{\rm sph}$ correlation is also expected from other two well known scaling relations 
involving the spheroid luminosity, $L_{\rm sph}$: 
the $L_{\rm sph} - n_{\rm sph}$ and the $M_{\rm BH} - L_{\rm sph}$ correlations.

However, obtaining an accurate estimate of the spheroid S\'ersic index requires a careful modelling of the galaxy's light distribution 
and some studies have failed to recover a statistically significant $M_{\rm BH} - n_{\rm sph}$ correlation.

With the aim of re-investigating the $M_{\rm BH} - n_{\rm sph}$ 
and other black hole mass scaling relations, 
we performed a detailed (i.e.~bulge, disks, bars, spiral arms, rings, halo, nucleus, etc.) 
decomposition of 66 galaxies, with directly measured black hole masses, 
that had been imaged at $3.6\rm~\mu m$ with \emph{Spitzer}. 
In this paper, the third of its series, we present an analysis of the $L_{\rm sph} - n_{\rm sph}$ and $M_{\rm BH} - n_{\rm sph}$ diagrams. 
While early-type (elliptical+lenticular) and late-type (spiral) galaxies split into two separate relations 
in the $L_{\rm sph} - n_{\rm sph}$ and $M_{\rm BH} - L_{\rm sph}$ diagrams, 
they reunite into a single $M_{\rm BH} \propto n_{\rm sph}^{3.39 \pm 0.15}$ sequence
with relatively small intrinsic scatter ($\epsilon_{(M_{\rm BH} | n_{\rm sph})} \simeq 0.25 \rm~dex$). 
{\bf to be finished...}
\end{abstract}

@@

and the relation between the central black hole mass, $M_{\rm BH}$, and the spheroid luminosity 
(e.g.~\citealt{kormendyrichstone1995,magorrian1998,marconihunt2003,haringrix2004}), 

@@

Motivated by the need for more accurate galaxy decompositions to refine and re-investigate scaling relations 

@@

\section{Analysis and results}
\label{sec:anal}
As in \emph{Paper II}, a linear regression analysis of the $L_{\rm sph} - n_{\rm sph}$ 
(Table \ref{tab:lregLn} and Figure \ref{fig:magn})
and $M_{\rm BH} - n_{\rm sph}$ (Table \ref{tab:lregMn} and Figure \ref{fig:mbhn}) diagrams 
was performed using three different routines: 
the BCES code from \cite{akritasbershady1996}, 
the FITEXY routine \citep{press1992}, as modified by \cite{tremaine2002}, 
and the Bayesian estimator {\tt linmix\_err} \citep{linmixerr}.
All of these three routines take into account the intrinsic scatter, 
but only the FITEXY and the {\tt linmix\_err} codes allow one to quantify it.
\cite{tremaine2002} cautioned that 
the BCES estimator becomes ineffective when the dataset contains at least one low-precision measurement -- 
regardless of how many high-precision measurements are in the sample -- 
and tends to be biased in case of low number statistics 
or if the mean square of the uncertainties associated to the independent variable is comparable to 
the variance of the distribution of the independent variable. 
According to the results from the Monte Carlo Markov Chain simulations of \cite{tremaine2002} and \cite{novak2006}, 
these problems can be overcome with the use of the modified FITEXY routine. 
\cite{park2012} also concluded that the modified FITEXY routine performs better and returns less biased results than the BCES estimator, 
and noted that the the modified FITEXY routine is computationally less intensive than the Bayesian technique {\tt linmix\_err}. 
Given that at least one of our subsamples (the lenticular galaxies) has a small size 
and that the uncertainties associated to $n_{\rm sph}$ are relatively large 
compared to the range spanned by the $n_{\rm sph}$ values for most of our subsamples, 
we put more trust in the results obtained with the modified FITEXY routine 
and throughout the text we quote only those. \\ 
We report both symmetrical and nonsymmetrical linear regressions.  
Symmetrical regressions are meant to be compared with theoretical expectations, 
whereas nonsymmetrical forward ($Y|X$) regressions -- which minimize the scatter in the $Y$ direction -- 
allow one to predict the value of the observable $Y$ with the best possible precision. \\
We searched for extreme outliers in both the $L_{\rm sph} - n_{\rm sph}$ and $M_{\rm BH} - n_{\rm sph}$ diagrams, 
and found that in our $L_{\rm sph} - n_{\rm sph}$ plot there are no $3\sigma$ outliers, 
whereas in our $M_{\rm BH} - n_{\rm sph}$ plot 
the lenticular galaxies NGC 0524 and NGC 3998 reside more than $3\sigma$ from the bisector linear regression for all galaxies. 
These two galaxies have therefore been excluded from the rest of the analysis.

@@

\section{Conclusions} 
The S\'ersic index of a galaxy's spheroidal component, $n_{\rm sph}$, constitutes a direct measure 
of its central radial concentration of stars. 
After \cite{graham2001bulges} proved that the central black hole mass, $M_{\rm BH}$, 
correlates with the stellar light concentration of a galaxy's spheroidal component, 
\cite{grahamdriver2007} presented and analyzed for the first time a tight $M_{\rm BH} - n_{\rm sph}$ correlation 
using a sample of 27 elliptical and disk galaxies for which they had performed photometric bulge/disk decomposition. 
The $M_{\rm BH} - n_{\rm sph}$ correlation can be predicted from the combination of two well known scaling relations 
involving the spheroid luminosity, $L_{\rm sph}$: 
the $M_{\rm BH} - L_{\rm sph}$ (e.g.~\citealt{kormendyrichstone1995,magorrian1998,marconihunt2003,haringrix2004}) 
and the $L_{\rm sph} - n_{\rm sph}$ (e.g.~\citealt{youngcurrie1994,jerjen2000,grahamguzman2003}). 
However, upon independently attempting photometric multicomponent decompositions 
for galaxy samples that were similar to that used by \cite{grahamdriver2007}, 
three subsequent studies \citep{sani2011,vika2012,beifiori2012} failed to recover a statistically significant $M_{\rm BH} - n_{\rm sph}$ relation. 
\cite{savorgnan2013} collected and compared the S\'ersic index measurements obtained by \cite{grahamdriver2007}, 
\cite{sani2011}, \cite{vika2012} and \cite{beifiori2012}, 
and argued that the discrepancies were due to inaccurate galaxy decompositions. \\
Moved by a urgent need to re-investigate and refine several black hole mass scaling relations, 
we performed state-of-the-art photometric multicomponent decompositions 
(i.e.~bulge, disks, bars, spiral arms, rings, halo, nucleus, depleted core, etc.) 
for 66 galaxies with a dynamical measurement of their black hole mass (\emph{Paper I}). 
We carefully measured the S\'ersic index of each galaxy's spheroidal component 
and estimated its associated uncertainty with a method that takes into account statistical and systematic errors. 
Our analysis shows that early- (elliptical + lenticular) and late-type (spiral) galaxies define two different correlations 
in the $L_{\rm sph} - n_{\rm sph}$ diagram, 
whereas they reunite in a single sequence in the $M_{\rm BH} - n_{\rm sph}$ diagram. 
With the current dataset, we measured an amount of intrinsic scatter in the $M_{\rm BH} - n_{\rm sph}$ diagram systematically smaller, 
but still consistent within the errors, with that observed in the the $M_{\rm BH} - L_{\rm sph}$ diagram (\emph{Paper II}). \\ 
%Although we refrain from the conclusion that the $M_{\rm BH} - n_{\rm sph}$ relation is more fundamental than the $M_{\rm BH} - L_{\rm sph}$, 
Our results suggest that the black hole mass is intimately connected to the spheroid central concentration of stars, 
which reflects the inner gradient of the spheroid gravitational potential. 
Besides conferring the spheroid S\'ersic index a predictive power to infer the black hole mass 
from a galaxy's image only (even photometrically uncalibrated), 
the $M_{\rm BH} - n_{\rm sph}$ correlation should become a fundamental ingredient in semi-analytic models and simulations of galaxy formation.

@@

%%
\acknowledgments
GS warmly thanks Alister Graham and Carlo Nipoti for useful discussion. 
%We also thank the anonymous referee for useful comments and suggestions. 
This research was supported by Australian Research Council funding through grants
DP110103509 and FT110100263.
This work is based on observations made with the IRAC instrument \citep{fazio2004IRAC} 
on-board the Spitzer Space Telescope, 
which is operated by the Jet Propulsion Laboratory, 
California Institute of Technology under a contract with NASA.
This research has made use of the GOLDMine database \citep{goldmine} and the NASA/IPAC Extragalactic Database (NED) 
which is operated by the Jet Propulsion Laboratory, California Institute of Technology, 
under contract with the National Aeronautics and Space Administration. 
We acknowledge the usage of the HyperLeda database (\url{http://leda.univ-lyon1.fr}).
The BCES routine \citep{akritasbershady1996} was run via the python module 
written by Rodrigo Nemmen \citep{nemmen2012}, which is available at \url{https://github.com/rsnemmen/BCES}.
